### 2014-11-26 fundamentals/impl.tex

\subsection{Cluster-Ready Software}

\subsection{Framework Structure}
% (OO, cpp, xml, backends, inheritances, mixins, generic apis, etc.)

\subsection{Dynamically Loadable Libraries}
% (diagram)

@@

\subsection{Agent Interchangeability}\label{sec:interchangeability}

@@

\subsection{Regions, Institutions, and Facilities}

@@

\subsection{Resources and Materials}
% Resources, Materials (note isotope tracking, decay behavior)

@@

\subsection{Toolkit}

@@

\subsubsection{Simulation Tools}

@@

\subsubsection{Nuclear Engineering Tools}

@@

\subsubsection{Toolkit Extensions}

@@

\subsection{Cycamore}
% base modules

@@

>>>>

\subsection{Modular Architecture}
% Motivation for encapsulated plug-in architecture 
% enables ecosystem, ease of contribution across institutions
% enables myriad levels of simulation fidelity
% enables FC analyses that can compare ‘apples to apples'

The architecture of \Cyclus allows developers 
to define nuclear fuel cycle agents and processes independent of the simulation 
logic. This encapsulated `plug-in'
design choice provides two major benefits. First, analysts across research institutions can take advantage of the 
simulation logic \gls{API} and archetype ecosystem when they apply \Cyclus to 
their specific problem. This 
mend-and-extend paradigm, common in modern software development, reduces 
maintenance overhead and improves code robustness \cite{citation_needed}. 
The researcher can focus on creating or customizing nuclear 
facility, institution, resource, and toolkit models within their specific area 
of technical expertise. This extensibility, magnified by an open source 
development process, frees the community of users to contribute desired 
features, independent from the team of \Cyclus core developers. 
% we could use a specific example of Cosi's equivalence method below
Second, because \Cyclus uses a modular archetype approach, comparing two 
archetypes is straightforward. Consider a case in which an analyst would like 
to compare the effect of using different models to determine the input fuel 
composition for fast reactors. A variety of fuel fabrication archetypes can be 
developed and interchanged while keeping the rest of the fuel cycle fixed
 to compare fuel fabrication designs.

\subsubsection{Dynamically Loadable Libraries}
% (diagram)

@@

\subsection{Agent-based Paradigm}
% superior detail in capturing simulation dynamics
% more flexible control over behavior
% describe Region/Institution/Facility hierarchy
% note importance of generic resource exchange paradigm

\Cyclus implements an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and it incorporates modern computing resources.  
Furthermore, \gls{ABM} provides a natural 
perspective from which to model a given problem. In the nuclear fuel cycle 
context, for example, an analyst can design an agent with reactor-related logic 
that is entirely independent from an agent with fuel fabrication-related 
logic. Defining the behavior of both agents according to the 
\gls{API} contract is sufficient for them to interact with one another as 
bona fide actors in the simulation.  The two archetype libraries can be used in the same
simulation without any shared knowledge.

Furthermore, the \gls{ABM} paradigm is easily comparable to system dynamics models.
System dynamics is a popular approach for modeling nuclear fuel cycles 
\cite{jacobson_vision_2009,van_den_durpel_daness_2009,guerin_impact_2009,guerin_benchmark_2009}. 
Formally however, system dynamics models are simply a strict subset of agent-based models
\cite{macal_agent-based_2010}.
That is, any system dynamics model can be translated
into an agent-based model. Furthermore, because agent-based
techniques can provide a richer model than can system dynamics, they enable the 
broadest range of simulations in a generic fashion.

\subsubsection{Agent Interchangeability}\label{sec:interchangeability}

@@

\subsubsection{Regions, Institutions, and Facilities}

@@

\subsection{Discrete Objects}
% enables more realistic models and metrics
% material routing metrics
% shadow fuel cycles
% etc.

\Cyclus models facilities, institutions, regions, and resources as discrete 
objects. A discrete resource model allows for a range of modeling granularity. In the
macroscopic extreme, it is equivalent to time-stepped continuous flow. In the
microscopic extreme, the model is capable of representing arbitrarily small 
material objects at isotopic resolution. In this way, \Cyclus is 
applicable across the full range of modeling extremes.

Although some questions require resolution at the level of 
individual facilities and materials, 
Fleet-based, lumped-material models do not distinguish between discrete
facility entities or materials.  As a result, many detailed performance 
metrics cannot be captured with previously existing tools.
The generality of \Cyclus therefore enables many use cases 
that cannot be achieved with fleet-based models.

One such use case seeks to capture system vulnerability to nefarious 
material diversion. Provenance and trade-history of distinct materials is the fundamental 
information unit in such studies, and so this type of analysis requires
 discrete simulation of a 
target facility and the individual materials modified within it. 
Provenance and trade-history of distinct materials is the fundamental 
information unit in such studies. 
Material risk analysis, therefore, demands discretely modeled objects like those 
in \Cyclus.

There are a number of types of users and developers who will 
interact with \Cyclus. National laboratory researchers,
academic researchers, private-industry economic forecasters, and even students 
in engineering courses may benefit from the ease of development and
customization provided by the discrete, modular, agent-based architecture of \Cyclus.

\subsubsection{Resources and Materials}
% Resources, Materials (note isotope tracking, decay behavior)

@@

\subsubsection{Dynamic Resource Exchange}

Placeholder as a reminder to add more detail about this.

\subsection{Software Decisions}

\subsubsection{Cluster-Ready Software}

\subsubsection{Framework Structure}
% (OO, cpp, xml, backends, inheritances, mixins, generic apis, etc.)

\subsection{Simulation Assistance}

\subsubsection{Toolkit}

@@

\paragraph{Simulation Tools}

@@

\paragraph{Nuclear Engineering Tools}

@@

\paragraph{Toolkit Extensions}

@@

\subsubsection{Cycamore}
% base modules

@@

\subsection{Quality Assurance}
% Organize this section according to major topics
% give each topic a section heading in boldface.
% try to cover the major common points :
%
% problem design
% methods of measurement
% supporting models
% supporting data
% simulations run
% results

% Just write the section headings for each part and indicate what goes in that
% section with words :
%
% heading
% figures (with captions)
% schematics (with captions and footnotes)
% equations
% tables

% What does it mean?
% What did I actually test?
% What were the results?
% Did the work yield a new method?
% Did the work yield new knowledge?
% What measurements did I make?
% How were these measurements characterized?
% What methods were used?
% What were the results?
% How were the measurements made and characterized?

Simulation science - like experimental science - has since the beginning had the 
problem of identifying trustworthy results from nonsensical results or noise.
This is epitomized by Charles Babbage, ``On two occasions I have been asked, 
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right 
answers come out?' ... I am not able rightly to apprehend the kind of confusion 
of ideas that could provoke such a question.'' \cite{babbage_passages_2011}. 
A simulator must both guard against the \emph{garbage in, garbage out} phenomenon, and also assure correctness.

Multiple strategies, collectively known as \emph{\gls{QA}}, have 
been invented over the years to mitigate the structural and algorithmic errors
on the part of simulators themselves. These include \emph{\gls{VV}}
\cite{boehm_software_1989}, \emph{\gls{UQ}} 
\cite{sacks_design_1989}, testing, and others. \gls{VV} and \gls{UQ} come from 
a distinctly computational science background while testing \emph{et al.} come from software development.

Nuclear engineering code quality is often governed by \gls{NQA1}, an 
\gls{ASME} specification 
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}.  
Although \gls{NQA1} is primarily 
used for designing reactors, it is general enough to apply to 
\Cyclus. \Cyclus has adopted an \emph{agile} development process 
\cite{larman_agile_2004}, 
interpreting \gls{NQA1} in a manner similar to the process adopted by the 
\gls{DOE} within \gls{NEAMS} \cite{neams_nuclear_2013} or the 
PyNE toolkit \cite{biondo_quality_2014}.

Quality assurance is an on-going process throughout the 
entire life of the code. As a simulator where most modeling decisions are made 
by third party archetype developers or users, the most important feature of 
\gls{QA} is verification. Validation may be impossible and \gls{UQ} is beyond 
the scope of the kernel.

Verification may be defined as the question, ``Is \Cyclus being built correctly?'' 
To answer this question, we turn to the software development of notions of testing,
documentation, version control, style guidelines, and continuous integration. 
This suite of process controls supplies mechanism for reliable and reproducible 
software. The impetus to implement these controls correctly is particularly strong in a scientific 
context because of the emphasis on reproducibility and provenance. These verification techniques are 
discussed individually in the sections that follow.

Validation, in contrast,  may be defined as the question, 
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal 
validation is not possible. Nonetheless, code-to-code comparisons with fuel cycle
simulators that use other modeling paradigms are underway, if nascent. However, such 
exercises are more likely to bring into relief the differences between the modeling
paradigms than supply substantial \gls{QA} and validation.

Lastly, \gls{UQ} is a process that is used on specific simulation
instantiations to statistically determine quality. \gls{UQ} is a process that applies 
to the running of many \Cyclus simulations. The kernel itself, as a generic fuel cycle 
simulator, cannot make reliable statements about the uncertainty of a particular fuel 
cycle. Any \gls{UQ} results are applicable only to the scenario that is under 
analysis by the user. \Cyclus is not a panacea for due dillegence on behalf of the fuel 
cycle modeler. Thus, \gls{UQ} is envisioned to be beyond the scope of core development 
and in the domain of the expert user.

Sections \ref{sec:qa-testing}-\ref{sec:qa-ci} discuss in greater detail the software 
development components that comprise the \Cyclus verification strategy.
Each of these on its own is a valuable addition to \gls{QA} but is cannot be the 
entire answer to the requirements imposed by \gls{NQA1}. Taken together and strictly 
adhered to, they present a fortress to protect against poorly designed or otherwise undesirable code.

\subsubsection{Testing}
\label{sec:qa-testing}

Automated software \emph{testing} is the first line of defense against
errors in implementation. It also acts as an early warning sign if the
simulator does not work as intended on a new system.
Testing serves a critical role in \gls{QA} because it directly compares the 
results of running software versus the expected behavior of the software.
In \Cyclus, three categories of tests are defined: unit tests, integration 
tests, and regression tests.

It is important to note that before a proposed
code change is allowed into \Cyclus,  the change must be covered by a test, either new or existing, and all of the tests must pass.  This is a
matter of policy for the developers. Of course, test code is 
just as likely to contain errors as the code it is designed to test. 
However, it is assumed that diligence and one level of testing 
is sufficient to meet the aims of \gls{QA} and prevent an infinite recursion
of testing-the-tests.

\paragraph{Unit Tests}

Unit tests compare the results of the smallest code \emph{unit}, 
typically a single function or a class. What constitutes the smallest code
unit depends on the context of the specific unit in question. While this is 
an ambiguous definition, when combined with a philosophy of keeping units as small
as possible, it is often clear in practice what needs to be tested.

\Cyclus uses the Google Test framework \cite{inc_googletest_2008} as a harness for running unit 
tests. Sufficient unit tests are required for any proposed change to the \Cyclus
code base. Currently \Cyclus implements over 450 unit tests and \Cycamore implements 
85.  These cover approximately 65\% of their respective code bases, and their numbers are expected to grow over time.

\paragraph{Integration Tests}

Integration tests combine multiple elements of the 
\Cyclus interface and test that they work correctly with each other.  By analogy, 
simply because the gears (units) are made correctly does not imply that the 
clock (integration) will run smoothly, run at all, or give the correct time.
In \Cyclus and \Cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of 
time. A set of standard input files have thus been constructed. 
Python is used to run the same set of \Cyclus simulations as a subprocess for each integration test.  The results are then inspected and compared via Nose \cite{pellerin_nose_2007}, a Python test framework.
In this way, \Cyclus code units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that 
major \Cyclus components are functioning as expected.

\paragraph{Regression Tests}

Regression tests ensure significant changes do not 
occur over the course of \Cyclus development. Such a change is called a 
\emph{regression} because the new version is almost always wrong.
Regression tests are implemented similarly to integration tests.
Nose is used to execute full \Cyclus simulations whose results
are then compared. In this category, however, the comparison is done against 
the output of the same input file when run with a previous version of \Cyclus, 
typically the last released version.
In some sense, regression tests are `dumb' in that they do 
not care about the contents of a simulation being correct, only whether or not 
it changed. They pick up glaring time-sensitive mistakes that may 
not appear elsewhere, though they do not purport to make claims about the
quality of the physics modeled. Regression tests are primarily implemented
in \Cycamore because its archetypes are sophisticated enough
to make interesting input files. Only a handful ($<10$) of time steps are needed 
for regression testing.

\subsubsection{Documentation}

All of the public interface (the \gls{API}) must be documented as required by the \Cyclus
\gls{QA} policy. This certifies that the intention for a code unit is communicated 
in way that is separate from its implementation. This policy is important 
because implementations are often wrong and do not reliably communicate their intent. Truly trustworthy communication about what 
software should be doing occurs when the author explicitly states the 
goals of an \gls{API} in prose form. Future development of the code is then able to 
compare the implementation versus the documentation to discern whether or not they 
are in agreement. Documentation is thus a secondary higher-level information 
stream.  In \Cyclus, this information is aggregated together into static 
websites with the Doxygen \cite{van_heesch_doxygen:_2008} and Sphinx 
\cite{brandl_sphinx_2014} 
tools, and can be accessed at \url{http://fuelcycle.org}.

\subsubsection{Version Control}

Version control is the history-preserving mechanism by which the provenance of 
a code is recorded. \Cyclus uses a well-established tool called git 
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utility in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every 
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there are many praises and criticisms of git, one 
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simultaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an 
accompanying message. Thus 
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.

\Cyclus uses a strategy known as \emph{git flow} 
\cite{kalliamvakou_code-centric_2014} 
to manage topical branches where individuals develop code: the mainline development
branch (simply known as develop), and the stable branch (called master).
In addition to these rules and delineations, \Cyclus also imposes rules on 
when it is acceptable from a quality control standpoint to merge code from 
the personal topical branches into the develop branch. All software at this stage 
must have sufficient successful tests and comprehensive documentation. Without
these core pieces, proposed changes are not allowed into \Cyclus. As an example, a schematic of 
the development stages between reporting a bug and merging the fix is shown in 
Figure \ref{fig:gitprocess}.

\begin{figure}[htbp]
\begin{center}
\includegraphics[height=0.6\textheight]{./images/gitprocess}
\end{center}
\caption{A version-control enabled process for finding, reporting, and fixing a bug 
can take place within an issue tracker alongside multiple stages of code 
testing and review. In the \Cyclus case, the final review stage includes unit, 
regression, and integration testing as well as manual code inspection.}
\label{fig:gitprocess}
\end{figure}

The act of proposing a change is known as a \emph{pull request}. The main \Cyclus repository is 
hosted remotely on the GitHub website \cite{dabbish_social_2012}. The online
mechanism for pull requests allows for code review by a member of the \Cyclus 
core team, separate from the author, prior to any change being included. Non-members
of the \Cyclus development team are allowed and encouraged to submit and review 
pull requests. However, only members of the \Cyclus core team are allowed to 
merge in any changes and only after the \gls{QA} standards have been met.

\subsubsection{Style Guide}

In any multi-person software project, there is a tension between how individuals
wish to write code. Every person tends to have their own custom style. To remedy this,
coding style guides are the software analogy to those for written language,
such as the Chicago Manual of Style. \Cyclus strictly enforces the use of the 
Google C++ Style Guide \cite{weinberger_google_2008} for all software contributions.
This means that all developers of \Cyclus nominally write \Cyclus code in the same 
way.  This homogenization may be a hurdle to new developers but ultimately 
improves code legibility. Adoption of a single style guide also settles
timeless arguments over what style is `the best.' Having only one means that 
there is a common enemy for all developers to dislike equally.

\subsubsection{Continuous Integration}
\label{sec:qa-ci}

\emph{\acrfull{CI}} is the idea that software should be tested and validated 
as it is being developed, rather than as a final stage in a longer development 
cycle.  Under \gls{CI}, every pull request is tested independently immediately after 
it is proposed. Such tests are run on all officially supported platforms. 
\Cyclus uses a \gls{CI} platform called Polyphemus \cite{scopatz_polyphemus_2014}. Polyphemus serves as an intermediary between GitHub pull requests on the frontend 
and temporary \Cyclus servers on the backend. These servers are hosted by 
the Build \& Test Laboratory (BaTLab) \cite{uw_batlab_team_batlab_2014} at the University of 
Wisconsin-Madison. Whenever a pull request is created, Polyphemus performs 
the following actions:

\begin{enumerate}
    \item Copies \Cyclus with the pull requested code to BaTLab,
    \item Initializes Linux and Mac OSX servers,
    \item Builds \Cyclus on all platforms,
    \item Runs the complete \Cyclus test suite, 
    \item Reports back whether the above steps succeeded or not.
\end{enumerate}

Since these steps are performed for all incoming code, it is easy for the 
\Cyclus core team to determine whether or not an individual pull request 
actually works. This helps identify build and test problems prior to 
broken code being allowed into the develop branch. \gls{CI} is a necessary 
but not sufficient addition to the \Cyclus \gls{QA}
system. Its merit is that it prevents bad code from being implemented. However, \Cyclus will always
require human eyes and hands to let good code in.
